---
title             : "The title"
shorttitle        : "Title"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

author_note: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  Enter abstract here. Each new line herein must be indented, like this line.
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
mask              : no

class             : "man"
output            : papaja::apa6_pdf
---

```{r load_packages, include = FALSE}
library("papaja")
```

Similar style introduction as the bias1 paper:
- p values are bad and we shouldn't do it
- NHST has it's flaws
- psychology's glorious new era

effect size
- these are a good idea
- what are variance overlap statistics
- talk about the newer statistics and why we use them (eta, omega, ges families)

limitations on effect size
- they are biased!
- that bias has consequences for sample size planning and such

current study

# Method

## Simulations
talk about all the variables and stuff here
formulas
manipulated variables = 
- sample size
- design type
- effect size type
- effect size size
- correlation
- levels

# Results

I picked the variables because we already are pretty clear of the effects of sample size and then I figured we could graph each of the 5 effect sizes across sample size with a three part cowplot for small medium and large effects. 

## Bias of Effects

### Levels

talk here about bias by number of levels examining the bias X effect size type X effect size across sample size 

### Design Type

talk here about bias by the type of design examining the bias X effect size type X effect size across sample size

### Correlation specifically RM since BN doesn't matter

talk here about bias by correlation examining the bias X effect size type X effect size across sample size 

## Power and Sample Size Planning

within each crazy way combination we figured out bias, then calculated the difference in sample size blah blah blah.

### Levels

### Design Type

### Correlation 

# Discussion


\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
